Sex-Specific Evolution of the Meiotic Recombination Rate

### 2020-05-21 doc/MS_v2.Rmd
DEBUGGED and CLEANED up the setup rmd and MS_v2 scripts

library(raster)
library(lme4)
library(nlme)
library(RLRsim)
#library(ggpubr)
library(reshape2)
library(patchwork)
library(png)
library(grid)
library(gridBase)
library(gridExtra)
library(dplyr)
library(cowplot)
library(kableExtra)
library(xlsx)

@@

load(file="~./MLH1repo/data/Results_setup_5.14.20.RData") #Results_setup_5.14.20.RData

@@

We used a panel of wild-derived inbred strains of house mice (*Mus musculus*) and related murids to profile natural genetic variation in recombination (Table 1). Our survey included 5 strains from **M. m. musculus**, 4 strains from _M. m. domesticus_, 2 strains from _M. m. molossinus_, 2 strains from **M. m. castaneus**, and 1 strain each from **M. caroli**, **M. spicilegus**, and **M. spretus**. We subsequently denote strains by their abbreviated subspecies and name (e.g. **domesticus^WSB^**).  Mice were housed at dedicated, temperature controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine. Mice from an inbred strain of Gough Island mice were sampled after XX generations of brother-sister mating. All mice were provided with ad libitum food and water. All procedures followed protocols approved by IACUC.

@@

*new* - The same dry-down spread technique was applied to both spermatocytes and oocytes based on [@peters_1997], with adjustment for volumes. Spermatocyte spreads were collected and prepared as described in [@peterson2019]. The majority of mice used MLH1 quantification were between 5 and 12 weeks of age. Juvenile mice between 12 and 15 days of age were used for DMC1 quantification. Both ovaries were collected from embryos (16-21 embryonic days) or neonates (0-48 hours after birth). Whole testes were incubated in 3ml of hypotonic solution for 45 minutes. Decapsulated ovaries were incubated in 300ul of hypotonic solution for 45 minutes. Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution. Cells were fixed in this solution and dried in a humid chamber at room temperature overnight. The following morning, slides were treated with a Photoflow wash (Kodak, cite). Slides were stored at -20 \* C if not stained immediately.
To visualize the structure of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3). Crossovers (COs) were visualized as MLH1 foci. Double strand breaks (DSBs) were visualized as DMC1 foci. The staining protocol followed [@anderson1999] and [@koehler2002]. Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37\* C. The master mix recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB. Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37\* C. Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB. The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37\* C. Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.

@@

To measure bivalent SC length, two image analysis algorithms were used. The first algorithm estimates the total (summed) SC length across bivalents for individual cells [@wang2019_sc]. The second algorithm estimates the SC length of individual bivalents [@peterson2020]. Both algorithms apply a ‘skeletonizing’ transformation to synapsed chromosomes that produces a single, pixel-wide ‘trace’ of the bivalent shape.

Total SC length per cell was quantified from pachtyene cell images [@wang2019_sc]. To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set. Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (MLH1.merge.Skel %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` -- 3,371 out of `r formatC( length(  (MLH1_data %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")`-- 4,143 cells with MLH1 counts.

SC length of individual bivalents was quantified in pachytene cell images [@peterson2020]. The DNA CrossOver algorithm [@peterson2020] isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci. The algorithm substantially speeds the accurate measurement of bivalents, but it sometimes interprets overlapping bivalents as single bivalents. In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 ( _molossinus^MSM^_ male) to`r max(algorithm.stat.category$mean_pass.rate)` -- 0.72 ( _musculus^KAZ^_ female). From the total set of cell images, `r formatC( length(unique(new.Curate.FULL.org %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` -- 10,458 bivalent objects were isolated by the algorithm. Following a manual curation, `r formatC( length(unique(new.Curate.FULL.org %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` -- 9,829 single-bivalent observations remained. The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2020]. The curated single bivalent data supplements our cell wide MLH1 count data with MLH1 foci counts for single bivalents. The proportions of bivalents with the same number of MLH1 foci where compared across strains by a chi-squared test using prop.test().

@@

@@

male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+

@@

```{r MLH1.cell.examples, echo=FALSE}

@@

#Cell GGPLOT
library(ggpubr)

@@

#try cowplot for more control of sizing
library("cowplot")

@@

We next fit models separately for  `r length(fomo.male_mouse.av$mouse)` -- 192 males and `r length(female_mouse.avs_for.analysis$mouse)` -- 144 females (M4). In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = 6.3110^{-8} -- `r sum_MLH1.male.glm.M4$coefficients[,4][8]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[8]` -- 6.1 foci), and _musculus^SKIVE^_ (p = 0--`r sum_MLH1.male.glm.M4$coefficients[,4][10]` ; effect = 3.8-- `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p=2.4210^{-12} -- `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]` -- 7.0).

@@

```{r male.strain.table, echo=FALSE}
M.strain.table <- ddply(fomo.male_mouse.av, c("strain"), summarise,
                        Nmice =length(unique(mouse)),

@@

glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
sum.glmMLH1Var_M3 <- summary(glmMLH1Var_M3)
#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)

@@

```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
#total-SC
#Q1.lmer_totSC.mouse.av.table -- #can't find this table
#Q1.lmer_totSC.mouse.av.table <- mouse.level.skel %>%

Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)

#subsp effect, many strain, #male
#G * male v.sig and male*LEW

Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
#LOT OF SIGNIFICant strain effects in this model
```

@@

summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)

@@

MLH1.merge.Skel$quality <- as.numeric(MLH1.merge.Skel$quality)

totSC_Q12_mouse.av.table <- ddply(.data= MLH1.merge.Skel %>% filter(quality < 3), 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),

@@

#REMOVE JUVI mice
tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]
totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice, ]

totSC_Q12_strain.av.table <- ddply(.data=MLH1.merge.Skel %>% filter(quality < 3), 
                 .(strain, sex),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),

@@

```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}

Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \n all bivalents")

@@

Q1_1CO.pos_t.test.FULL <- t.test((mouse.avs_4MM %>% filter(sex == "male") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") )$norm.first.foci.pos)

#WSB
Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
#G
Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
#LEW
Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)

#PWD
Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)

#KAZ
Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)

#SKIVE
#Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% #filter(strain == "SKIVE") )$norm.first.foci.pos,
 #          (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == #"SKIVE") )$norm.first.foci.pos )

#MSM
Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)

@@

#FULL all strains
Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,

                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
#p-value = 0.07

Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
#p-value = 1e-12

#WSB
Q1.IFD.raw_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
                                (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_WSB <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)

#LEW
Q1.IFD.raw_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_LEW <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)

#G
Q1.IFD.raw_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_G <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)

#PWD
Q1.IFD.raw_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_PWD <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)

#KAZ
Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
                                ( mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)

@@

glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM)
summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)

>>>>

library(raster)
library(lme4)
library(nlme)
library(RLRsim)
library(ggpubr)
library(reshape2)
library(patchwork)
library(png)
library(grid)
library(gridBase)
library(gridExtra)
library(dplyr)
library(cowplot)
library(kableExtra)
library(xlsx)

@@

load(file="~./MLH1repo/data/Results_setup_5.21.20.RData") #Results_setup_5.14.20.RData

#load(file="~./MLH1repo/data/Results_setup_5.14.20.RData") #Results_setup_5.14.20.RData

@@

We used a panel of wild-derived inbred strains of house mice ( _Mus musculus_ ) and related murids to profile natural genetic variation in recombination (Table 1). Our survey included 5 strains from _M. m. musculus_, 4 strains from _M. m. domesticus_, 2 strains from _M. m. molossinus_, 2 strains from _M. m. castaneus_, and 1 strain each from _M. caroli_, _M. spicilegus_, and _M. spretus_. We subsequently denote strains by their abbreviated subspecies and name (e.g. _domesticus^WSB^_).  Mice were housed at dedicated, temperature controlled facilities in the UW-Madison School of Medicine and Public Health, with the exception of mice from Gough Island, which were housed in a temperature-controlled facility in the UW-Madison School of Veterinary Medicine. Mice from an inbred strain of Gough Island mice were sampled after XX generations of brother-sister mating. All mice were provided with ad libitum food and water. All procedures followed protocols approved by IACUC.

@@

*new* - The same dry-down spread technique was applied to both spermatocytes and oocytes based on (@peters_1997), with adjustment for volumes. Spermatocyte spreads were collected and prepared as described in [@peterson2019]. The majority of mice used MLH1 quantification were between 5 and 12 weeks of age. Juvenile mice between 12 and 15 days of age were used for DMC1 quantification. Both ovaries were collected from embryos (16-21 embryonic days) or neonates (0-48 hours after birth). Whole testes were incubated in 3ml of hypotonic solution for 45 minutes. Decapsulated ovaries were incubated in 300ul of hypotonic solution for 45 minutes. Fifteen microliters of cell slurry (masticated gonads) were transferred to 80ul of 2% PFA solution. Cells were fixed in this solution and dried in a humid chamber at room temperature overnight. The following morning, slides were treated with a Photoflow wash (Kodak, cite). Slides were stored at -20 \* C if not stained immediately.
To visualize the structure of meiotic chromosomes we used antibody markers for the centromere (CREST) and lateral element of the synaptonemal complex (SC) (SYCP3). Crossovers (COs) were visualized as MLH1 foci. Double strand breaks (DSBs) were visualized as DMC1 foci. The staining protocol followed [@anderson1999] and [@koehler2002]. Antibody staining and slide blocking were performed in 1X antibody dilution buffer (ADB) (normal donkey serum (Jackson ImmnuoResearch), 1X PBS, bovine serum albumin (Sigma), and Triton X-100 (Sigma)).
Following a 30-minute blocking wash in ABD, each slide was incubated with 60ul of a primary antibody master mix for 48 hours at 37\* C. The master mix recipe contained polyclonal anti-rabbit anti-MLH1 (Calbiochem; diluted 1:50) or anti rabbit anti-DMC1) (mix of DMC1), anti-goat polyclonal anti-SYCP3, (Abcam; diluted 1:50), and anti-human polyclonal antibody to CREST (Antibodies, Inc; diluted 1:200) suspended in ADB. Slides were washed twice in 50ml ADB before the first round of secondary antibody incubation for 12 hours at 37\* C. Alexa Fluor 488 donkey anti-rabbit IgG (Invitrgoen, location; diluted to 1:100) and Coumarin AMCA donkey anti-human IgG (Jackson ImmunoResearch; diluted to 1:200) were suspended in ADB. The last incubation of Alexa Fluor 568 donkey anti-goat (Invitrogen; diluted 1:100) was incubated at 1:100 for 2 hours at 37\* C. Slides were fixed with Prolong Gold Antifade (Invitrogen) for 24 hours after a final wash in 1x PBS.

@@

To measure bivalent SC length, two image analysis algorithms were used. The first algorithm estimates the total (summed) SC length across bivalents for individual cells (@wang2019_SC). The second algorithm estimates the SC length of individual bivalents (@peterson2019). Both algorithms apply a ‘skeletonizing’ transformation to synapsed chromosomes that produces a single, pixel-wide ‘trace’ of the bivalent shape.

Total SC length per cell was quantified from pachtyene cell images [@wang2019_SC]. To reduce algorithmic errors in SC isolation, outliers were visually identified at the mouse level and removed from the data set. Mouse averages were calculated from cell-wide total SC lengths in `r formatC( length(  (Total.SC_DF  %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")` -- 3,371 out of `r formatC( length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name), format="d", big.mark=",")`-- 4,143 cells with MLH1 counts.

SC length of individual bivalents was quantified in pachytene cell images [@peterson2019]. The DNA CrossOver algorithm [@peterson2019] isolates single, straightened bivalent shapes, returning SC length, location of MLH1 foci, and location of CREST (centromere) foci. The algorithm substantially speeds the accurate measurement of bivalents, but it sometimes interprets overlapping bivalents as single bivalents. In our data set, average proportions of bivalents per cell isolated by the algorithm ranged from `r min(algorithm.stat.category$mean_pass.rate)` -- 0.51 ( _molossinus^MSM^_ male) to`r max(algorithm.stat.category$mean_pass.rate)` -- 0.72 ( _musculus^KAZ^_ female). From the total set of cell images, `r formatC( length(unique(cleanCurated_BivData %>% filter(SC.pass == "1"))$fileName ), format="d", big.mark=",")` -- 10,458 bivalent objects were isolated by the algorithm. Following a manual curation, `r formatC( length(unique(cleanCurated_BivData %>% filter(hand.foci.count != ""))$Obj.ID ), format="d", big.mark=",")` -- 9,829 single-bivalent observations remained. The accuracy of the algorithm is high compared to hand measures after this curation step [@peterson2019]. The curated single bivalent data supplements our cell wide MLH1 count data with MLH1 foci counts for single bivalents. The proportions of bivalents with the same number of MLH1 foci where compared across strains by a chi-squared test using prop.test().

@@

#change the segements and text!!

@@

male.plot <- ggplot( male_mouse.avs_for.analysis, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("")+scale_color_manual(values=colors_of_strains)+ylim(c(17,34))+

@@

```{r MLH1.cell.examples, echo=FALSE, message=FALSE, warning=FALSE}

@@

#Cell GGPLOT

@@

@@

We next fit models separately for  `r length(male_mouse.avs_for.analysis$mouse)` -- 192 males and `r length(female_mouse.avs_for.analysis$mouse)` -- 144 females (M4). In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = 6.3110^{-8} -- `r sum_MLH1.male.glm.M4$coefficients[,4][8]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[8]` -- 6.1 foci), and _musculus^SKIVE^_ (p = 0--`r sum_MLH1.male.glm.M4$coefficients[,4][10]` ; effect = 3.8-- `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p=2.4210^{-12} -- `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]` -- 7.0).

@@

```{r male.strain.table, echo=FALSE}

M.strain.table <- ddply(male_mouse.avs_for.analysis, c("strain"), summarise,
                        Nmice =length(unique(mouse)),

@@

glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)

@@

```{r Q1.glm_totalSC_M2.M3, echo=FALSE, message=FALSE}
#total-SC
#Q1.lmer_totSC.mouse.av.table -- #can't find this table
#Q1.lmer_totSC.mouse.av.table <- mouse.level.skel %>% 
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")

mouse.total.skel$subsp <- factor(mouse.total.skel$subsp, ordered = FALSE)
mouse.total.skel$sex <- factor(mouse.total.skel$sex, ordered = FALSE)
mouse.total.skel$strain <- factor(mouse.total.skel$strain, ordered = FALSE)
mouse.total.skel$species <- factor(mouse.total.skel$species, ordered = FALSE)

#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]

Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data= (mouse.total.skel[ ! mouse.total.skel$strain %in% remove_strains, ] %>% filter(species == "M.musculus") )  )
sum.Q1.glm_totSC_M2 <- summary(Q1.glm_totSC_M2)

Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=  (mouse.total.skel[ ! mouse.total.skel$strain %in% remove_strains, ] %>% filter(species == "M.musculus") ) )
sum.Q1.glm_totSC_M3 <- summary(Q1.glm_totSC_M3)
#LOT OF SIGNIFICant strain effects in this model
```

@@

summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)
```

@@

Total.SC_DF$quality <- as.numeric(Total.SC_DF$quality)

totSC_Q12_mouse.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),

@@

#REMOVE JUVI mice
tot.SC.Q12_juvi.male <- totSC_Q12_mouse.av.table[totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]
totSC_Q12_mouse.av.table <- totSC_Q12_mouse.av.table[! totSC_Q12_mouse.av.table$mouse  %in% juvi_mice.list, ]

totSC_Q12_strain.av.table <- ddply(.data= Total.SC_DF %>% filter(quality < 3), 
                 .(strain, sex),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),

@@

```{r Q1show.F1pos_Full, echo=FALSE, warning=FALSE}
#

Q1.mouse.av.biv_scatr_box   <- ggplot( (mouse.level.BivData_table %>% filter(species == "M.musculus") ), aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \n all bivalents")

@@

Q1_1CO.pos_t.test.FULL <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(species == "M.musculus") )$norm.first.foci.pos,
                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$norm.first.foci.pos)

#WSB
Q1_1CO.pos_t.test.WSB <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
#G
Q1_1CO.pos_t.test.G <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
#LEW
Q1_1CO.pos_t.test.LEW <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)

#PWD
Q1_1CO.pos_t.test.PWD <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)

#KAZ
Q1_1CO.pos_t.test.KAZ <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)

#MSM
Q1_1CO.pos_t.test.MSM <- t.test((mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
                           (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)

@@

#FULL all strains
Q1.IFD.raw_FULL_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" ) %>% filter(species == "M.musculus")  )$mean_IFD.2CO_ABS,

                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_IFD.2CO_ABS)
#p-value = 0.07

Q1.IFD.PER_FULL_t.test <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )  %>% filter(species == "M.musculus")    )$mean_IFD.2CO_PER,
                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(species == "M.musculus") )$mean_IFD.2CO_PER)
#p-value = 1e-12

#WSB
Q1.IFD.raw_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_ABS,
                                (mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_WSB <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="WSB" ))$mean_IFD.2CO_PER,
                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="WSB"))$mean_IFD.2CO_PER)

#LEW
Q1.IFD.raw_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW" ))$mean_IFD.2CO_ABS,
                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_LEW <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER,
                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="LEW"))$mean_IFD.2CO_PER)

#G
Q1.IFD.raw_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_ABS,
                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_G <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="G" ))$mean_IFD.2CO_PER,
                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="G"))$mean_IFD.2CO_PER)

#PWD
Q1.IFD.raw_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_ABS,
                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_PWD <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="PWD" ))$mean_IFD.2CO_PER,
                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="PWD"))$mean_IFD.2CO_PER)

#KAZ
Q1.IFD.raw_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_ABS,
                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_ABS)

Q1.IFD.PER_t.test_KAZ <- t.test( (mouse.level.BivData_table %>% filter(sex == "male") %>% filter(strain =="KAZ" ))$mean_IFD.2CO_PER,
                                ( mouse.level.BivData_table %>% filter(sex == "female") %>% filter(strain =="KAZ"))$mean_IFD.2CO_PER)

@@

glm.Q1_IFD.ABS.M2  <- glm(mean_IFD.2CO_ABS ~ subsp * sex * strain, data=Mouse.Table_BivData_4MM  %>% filter(species == "M.musculus")  )
summry.glm.Q1_IFD.ABS.M2 <- summary(glm.Q1_IFD.ABS.M2)
